---
title: | 
  | Econometrics III
  | Assignment Part 1 & 2
  | Tinbergen Insitute
author: |
  | Stanislav Avdeev \hspace{3em} Bas Machielsen 
  | 590050sa \hspace{5em} 590049bm 
  | stnavdeev@gmail.com \hspace{2em} 590049bm@eur.nl
date: \today
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  # Если мы не хотим код-чанков, можем менять на echo = FALSE
                      warning = FALSE,
                      message = FALSE,
                      out.width = "250pt", out.height = "200pt", 
                      fig.align = "center")

library(rio); library(tidyverse)
library(dynlm) ; library(stargazer)
library(normtest); library(reticulate)
library(AER)

```

<!-- I created a file .Renviron, where the python distribution is located on my system. You can find that by opening a terminal, and entering `$ which -a python python3`. Then, in RStudio, use `usethis::edit_r_environ()`, and add `RETICULATE_PYTHON="/Users/basmachielsen/opt/anaconda3/bin/python"` (or your directory) on a new line to the file. In this way, we can seamlessly interchange R and Python code chunks. Restart RStudio, and then everything is ready to go: --> 

## Question 1

**Part (a)**

```{r plotgdp}
# Part 1: Plot the Dutch GDP, ACF, and PACF
df <- readr::read_csv("./data/data_assign_p1.csv")

df <- ts(df$GDP_QGR, 
         frequency = 4, 
         start = c(1987, 2))
```

In the above figure, we observe the growth rates of Dutch GDP from 1987 to 2009. 

```{r plotacfpacf}
autoc <- acf(df, 
    lag.max = 12, plot = F) 

Box.test(df, 
         lag = 12, 
         type = "Ljung-Box")

pautoc <- pacf(df, 
     lag.max = 12, plot = F) 

plot(pautoc, main = "Partial Autocorrelation Function")
```

In the above figure, we plot the ACF and PACF, which is the ACF controlled for the other lagged correlations. The ACF tells us that the correlation of GDP growth with its lags is very low - hinting at very little time-dependence in this time-series. More precisely, the estimated correlation coefficients are not higher than 0.2 (for the lag of 1 period). If the GDP is indeed generated by an AR(p) process, the estimates show that the process has low $\phi$'s (in absolute value), indicating a low time dependence. 

**Part (b)**

```{r, results='asis'}
ar4 <- dynlm(df ~ L(df, 1) + L(df, 2) + L(df, 3) + L(df, 4))
ar3 <- dynlm(df ~ L(df, 1) + L(df, 3) + L(df, 4))
ar2 <- dynlm(df ~ L(df, 1) + L(df, 3))
ar1 <- dynlm(df ~ L(df, 1))

ar1_m <- forecast::Arima(df, c(1,0,0), method = "CSS")

stargazer(ar4, ar3, ar2, ar1, type = "latex", header = FALSE)

```

The table above shows us the coefficients of the AR(p) models with 1 to 4 lags included. Using a general-to-specific approach, we eliminate insignificant at the 0.05 level lags at each step. Column (1) shows that only the first lag is significant at the 0.1 level. The second lag has the lowest $\psi$ coefficient and the highest s.e., thus, we eliminate it first. As a result, we are left with the first lag in the model which is significant at the 0.05 level.

**Part c**

```{r}
# Part 3: Plot ACF of residuals
autocred <- acf(ar1_m$resid, 12, 
                plot = FALSE)
durbinWatsonTest(ar1)
```

```{r figure, echo = FALSE}
par(mfrow=c(2,2))

plot(df, 
     xlab = "Time", 
     ylab = "GDP Growth (%)", 
     main = "GDP Growth over Time")

plot(autoc, 
     main = "Autocorrelation Function")

plot(pautoc, 
     main = "Partial Autocorrelation Function")

plot(autocred, main = "Autocorrelation Function of Residuals", 
           xlab = "Time", 
           ylab = "Autocorrelation")

```

**Part d**


```{r}
# Part 4: Forecast AR model for 2 years
df_pred <- predict(ar1_m, n.ahead = 8)$pred

```

**Part e**

```{r}
# Part 5: Produce CI
df_ciu <- predict(ar1_m, n.ahead = 8)$pred + predict(ar1_m, n.ahead = 8)$se*1.96
df_cil <- predict(ar1_m, n.ahead = 8)$pred - predict(ar1_m, n.ahead = 8)$se*1.96
```

**Part f**

```{r}
# Part 6: Check normality
jb.norm.test(ar1_m$resid)
# reject H0, innovations are not normally distributed
```

## Question 2

```{r}
# Get the data
df <- readr::read_csv("./data/data_assign_p2.csv")

df <- ts(df[,2:3], 
         frequency = 4, 
         start = c(1987, 2))
```

**Part (a)**

```{r}
# Part a: Plot the Dutch GDP/unemployment, AR and ADL
plot(df)

ar4 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 2) + L(df[,1], 3) + L(df[,1], 4))
ar3 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 3) + L(df[,1], 4))
ar2 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 3))

adl4 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 2) + L(df[,1], 3) + L(df[,1], 4) + L(df[,2], 1) + L(df[,2], 2) + L(df[,2], 3) + L(df[,2], 4))
adl3 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 3) + L(df[,1], 4))
adl2 <- dynlm(df[,1] ~ L(df[,1], 1) + L(df[,1], 3))

stargazer(adl4, adl3, adl2, type = "text", header = FALSE)
```

