---
title: | 
  | Econometrics III
  | Assignment Part 1 & 2
  | Tinbergen Insitute
author: |
  | Stanislav Avdeev \hspace{3em} Bas Machielsen 
  | student no \hspace{5em} 590049bm 
  | stnavdeev@gmail.com \hspace{2em} 590049bm@eur.nl
date: \today
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 4, fig.height = 3)

library(rio)
library(tidyverse)
library(dynlm) # estimate AR model
library(stargazer)
library(normtest)
library(reticulate)

# dont use wd
```

## Instructions

I created a file .Renviron, where the python distribution is located on my system. You can find that by opening a terminal, and entering `$ which -a python python3`. Then, in RStudio, use `usethis::edit_r_environ()`, and add `RETICULATE_PYTHON="/Users/basmachielsen/opt/anaconda3/bin/python"` (or your directory) on a new line to the file. In this way, we can seamlessly interchange R and Python code chunks. Restart RStudio, and then everything is ready to go:

## Question 1

**Part (a)**

```{r}
# Part 1: Plot the Dutch GDP, ACF, and PACF
df <- readr::read_csv("./data/data_assign_p1.csv")
df <- ts(df$GDP_QGR, frequency = 4, start = c(1987, 2))

plot(df)
acf(df, lag.max = 12)
Box.test(df, lag = 12, type = "Ljung-Box")
pacf(df, lag.max = 12)
```

**Part (b)**

```{r, results='asis'}
ar4 <- dynlm(df ~ L(df, 1) + L(df, 2) + L(df, 3) + L(df, 4))
ar3 <- dynlm(df ~ L(df, 1) + L(df, 3) + L(df, 4))
ar2 <- dynlm(df ~ L(df, 1) + L(df, 3))
ar1 <- dynlm(df ~ L(df, 1))
stargazer(ar4, ar3, ar2, ar1, type = "latex")

ar1_m <- arima(df, c(1, 0, 0))
```

**Part c**

```{r}
# Part 3: Plot ACF of residuals
acf(ar1_m$resid, 12)
```

**Part d**

```{r}
# Part 4: Forecast AR model for 2 years
df_pred <- predict(ar1_m, n.ahead = 8)$pred
```

**Part e**

```{r}
# Part 5: Produce CI
df_ciu <- predict(ar1_m, n.ahead = 8)$pred + predict(ar1_m, n.ahead = 8)$se*1.96
df_cil <- predict(ar1_m, n.ahead = 8)$pred - predict(ar1_m, n.ahead = 8)$se*1.96
```

**Part f**

```{r}
# Part 6: Check normality
jb.norm.test(ar1_m$resid)
# reject H0, innovations are not normally distributed
```

```{python}
import pandas as pd
import numpy as np

r.mtcars.sum()
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Including Matplotlib

```{python}
hoi = np.arange(0,10)

hoi
```


```{r}
#py$hoi
#gert::git_branch_checkout("attempt_bas")
#gert::git_add(c("*"))
#gert::git_commit(message = "Automatic commit")
#gert::git_push()
```
