---
title: | 
  | Econometrics III
  | Assignment Part 1 & 2
  | Tinbergen Insitute
author: |
  | Stanislav Avdeev \hspace{3em} Bas Machielsen 
  | student no \hspace{5em} 590049bm 
  | stnavdeev@gmail.com \hspace{2em} 590049bm@eur.nl
date: \today
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(reticulate)
```

## Instructions

I created a file .Renviron, where the python distribution is located on my system. You can find that by opening a terminal, and entering `$ which -a python python3`. Then, in RStudio, use `usethis::edit_r_environ()`, and add `RETICULATE_PYTHON="/Users/basmachielsen/opt/anaconda3/bin/python"` (or your directory) on a new line to the file. In this way, we can seamlessly interchange R and Python code chunks. Restart RStudio, and then everything is ready to go:

```{r cars}
summary(cars)

mtcars <- mtcars
```


```{python}
import pandas as pd
import numpy as np

r.mtcars.sum()
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Including Matplotlib

```{python}
hoi = np.arange(0,10)

hoi
```


```{r}
py$hoi
gert::git_branch_checkout("attempt_bas")
gert::git_add(c("*"))
gert::git_commit(message = "Automatic commit")
gert::git_push()
```
