---
title: | 
  | Econometrics III
  | Assignment Part 3, 4, 5
  | Tinbergen Insitute
author: |
  | Stanislav Avdeev \hspace{3em} Bas Machielsen 
  | 590050sa \hspace{5em} 590049bm 
  | stnavdeev@gmail.com \hspace{2em} 590049bm@eur.nl
date: \today
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(echo = TRUE,  # Если мы не хотим код-чанков, можем менять на echo = FALSE
                      warning = FALSE,
                      message = FALSE,
                      size = "small",
                      out.width = "300pt", out.height = "200pt", 
                      fig.align = "center")

library(rio); library(tidyverse)
library(dynlm) ; library(stargazer)
library(normtest); library(reticulate)
library(AER)

```

<!-- I created a file .Renviron, where the python distribution is located on my system. You can find that by opening a terminal, and entering `$ which -a python python3`. Then, in RStudio, use `usethis::edit_r_environ()`, and add `RETICULATE_PYTHON="/Users/basmachielsen/opt/anaconda3/bin/python"` (or your directory) on a new line to the file. In this way, we can seamlessly interchange R and Python code chunks. Restart RStudio, and then everything is ready to go: --> 

## Question 3

**Part 1**:

First, we plot the two time series:

```{r plot}
df3 <- readr::read_csv("./data/data_assign_p3.csv")

twostocks <- c("apple", "netflix")

df3_twostocks <- df3 %>%
    janitor::clean_names() %>%
    pivot_longer(-date) %>%
    filter(is.element(name, twostocks)) %>%
    mutate(date = lubridate::dmy(date)) 

df3_twostocks %>%
    ggplot(aes(x = date, y = value, 
               group = name, color = name)) + geom_line()
```

Then, we show the acf and pacf-functions:

```{r autocor}

n1 <- df3$NETFLIX %>%
    acf(lag.max = 12, plot = F)

n2 <- df3$NETFLIX %>%
    pacf(lag.max = 12, plot = F)

a1 <- df3$APPLE %>%
    acf(lag.max = 12, plot = F)

a2 <- df3$APPLE %>%
    pacf(lag.max = 12, plot = F)

par(mfrow=c(2,2))

plot(n1, main = "Netflix");plot(n2, main = "Netflix")
plot(a1, main = "Apple"); plot(a2, main = "Apple")
     
```

The ACF's tell us that the stock price is highly dependent on the past stock price, and this dependence decays only very slowly: even the 50 or 100-period lag still shows significant autocorrelation. 

\clearpage

**Part 2**

We now implement a general to specific unit root test function:

```{r g2s function}
unit_root_test <- function(column, order, critical_value){
  
  # Make the dataset
  series <- ts(column) 
  first_differences <- diff(series, differences = 1)
  laggedvar <- stats::lag(series, -1)
  
  #other lagged first differences, delta x_{t-1}, .., delta x_{t-p+1}
  lagged_fds <- list()
  
  for(i in 1:(order-1)){
    
    lagged_fds[[i]] <- stats::lag(first_differences, k = -i)
    
  }
  
  df <- cbind(first_differences, laggedvar, purrr::reduce(lagged_fds, cbind)) %>%
    as_tibble()
 
  colnames(df) <- c("dxt", "xtm1", paste("dxtm", 1:(order-1), sep = ""))
  
  df <- df %>%
    na.omit()
  
  # run the stepwise regression - find the best model
  null = lm(data = df, formula = "dxt ~ xtm1")
  full = lm(data = df, formula = paste("dxt ~ xtm1 +", 
                                       paste(paste("dxtm", 1:(order-1), sep = ""), 
                                             collapse = ' + '),
                                       collapse = " ")
  )
  
  bestmodel <- step(full,
       scope = list(lower = null, upper = full),
       direction = "backward",
       criterion = "BIC",
       k = log(nrow(df)))
  
  # perform the unit root test (MacKinnon, 2010)
  b_critical <- critical_value
  
  t_value <- bestmodel %>%
    summary() %>%
    .$coefficients %>%
    .[,3] %>%
    .["xtm1"]
  
  significant = abs(t_value) > abs(b_critical)
    
  data.frame(stock = deparse(substitute(column)),
             best_model = as.character(bestmodel$call[2]), 
             t_value = t_value, 
             sig = significant)
}


```


```{r, message = FALSE, warning = FALSE, results = "hide"}
summary_tests <- list()

for(i in 1:length(colnames(df3[,-1]))){
  df <- unit_root_test(df3[, i+1], 12, -1.6156)
  summary_tests[[i]] <- df %>%
    mutate(name = colnames(df3[,-1][i]))
}

summary_tests <- summary_tests %>%
  purrr::reduce(rbind) %>%
  select(-stock)

rownames(summary_tests) <- NULL
```

```{r table}
knitr::kable(summary_tests)
```

The test seems to be significant for a number of stocks, indicating that for these stocks, the null hypothesis of a unit root is rejected in favor of stationarity. With a 10% $\alpha$-level, we expect to see a type I-error (rejecting the null while it is true) about one tenth of the time for every test. This means that the probability of having at least 1 type-I error is very large: $(1 - 0.9^{10})$ = `r 1-0.9^10`. It would be better to use some kind of Bonferroni correction to correct for these compounding type I-errors, but alternatively, we could also lower the $\alpha$-level. 

**Part 3**

The forecast $\mathbb{E}[P_{t+1} | D_t] = \mathbb{E}[P_{t+1} | P_t] = \mathbb{E}[P_t + \epsilon_t] = P_t$. Similarly, the forecast $\mathbb{E}[P_{t+2}] = \mathbb{E}[P_{t+1} + \epsilon_{t+1}] = \mathbb{E}[P_{t+1}] = P_t$. Generalizing this pattern, the forecast for $P_{t+h} = P_t$. The variance of the forecast is derived using the distribution:

\begin{align*}
P_{t+1} &= P_t + \epsilon_t \Rightarrow P_{t+1} | P_t \sim N(P_t, \sigma^2) \\
P_{t+2} &= P_{t+1} + \epsilon_{t+1} = \\
&= P_t + \epsilon_t + \epsilon_{t+1} \Rightarrow P_{t+2} | P_t \sim N(P_t, 2 \sigma^2) \\
\end{align*}

Generalizing this pattern, we can see that $\text{Var}(P_{t+h}) = h \cdot \sigma^2$. Hence, we can implement our forecasts in the following way:

```{r}
#forecast code
p_tplush <- df3 %>%
  slice_tail(n=1) %>%
  select(c("APPLE", "MICROSOFT")) 

var_apple <- lm(data = df3 %>%
     select("APPLE"), 
   formula = APPLE ~ lag(APPLE, 1)) %>%
  .$residuals %>%
  var()

var_microsoft <- lm(data = df3 %>%
     select("MICROSOFT"), 
   formula = MICROSOFT ~ lag(MICROSOFT, 1)) %>%
  .$residuals %>%
  var()

forecasts_apple <- data.frame(time = 1:5, 
                              value = rep(p_tplush %>%
                                            pull(1), 5),
                              var = var_apple * 1:5)

forecasts_microsoft <- data.frame(time = 1:5, 
                                  value = rep(p_tplush %>%
                                               pull(2), 5),
                                  var = var_microsoft * 1:5)

forecasts_apple %>%
  ggplot(aes(x = time)) + 
  geom_line(aes(y = value)) +
  geom_line(aes(y = value + 1.96*sqrt(var)), lty = "dashed") +
  geom_line(aes(y = value - 1.96*sqrt(var)), lty = "dashed")

forecasts_microsoft %>%
  ggplot(aes(x = time)) + 
  geom_line(aes(y = value)) +
  geom_line(aes(y = value + 1.96*sqrt(var)), lty = "dashed") +
  geom_line(aes(y = value - 1.96*sqrt(var)), lty = "dashed")

```

Hence, there is no investment advice that we can give: the value is predicted to remain constant, and for all predictions, the probability of an increase in stock price equals the probability of a decrease in stock price. The expected value of any investment strategy is 0, and the value is neither expected to increase, nor to decrease. 

**Part 4**

_Do  you  find  a  statistically  significant  contemporaneous  relation  between  Microsoft  and Exxon  Mobile  stock  prices?_

```{r, results='asis'}
lm(df3, formula = MICROSOFT ~ EXXON_MOBIL) %>%
  stargazer(header = F, omit.stat = c("adj.rsq", "ser", "f"))

```

_Do  you  agree  that  changes  in  Microsoft  stock  prices  are largely explained by fluctuations in the stock price of Exxon Mobile?_

We do not agree that changes in Microsoft stock prices are largely explained by fluctuations in the stock price of Exxon Mobile. It is likely that these results are driven by a shared stochastic trend in both of variables, in other words, the variables might be cointegrated. It can be shown that if two variables share a stochastic trend, the t-value of the estimated coefficient tends to infinity, and the probability of obtaining statistical significance to 1, even under the assumption of completely unrelated trends. 

## Question 4

**Part 1**

```{r}
df4 <- read_csv("data/data_assign_p4.csv")

df4$obs <- ts(df4$obs, 
              frequency = 4,
              start = c(1988, 1))

p_cons <- ggplot(data = df4, aes(x = as.Date(obs), y = CONS)) + 
  geom_line() + 
  scale_x_date(date_labels = "%b %Y")

p_inc <- ggplot(data = df4, aes(x = as.Date(obs), y = INC)) + 
  geom_line() + 
  scale_x_date(date_labels = "%b %Y")

cowplot::plot_grid(p_cons, p_inc)

inc <- ts(df4$INC,
   frequency = 4,
   start = c(1988, 1)) 

cons <- ts(df4$CONS,
   frequency = 4,
   start = c(1988, 1)) 

inc_acf <- acf(inc, plot = F); inc_pcf <- pacf(inc, plot = F)
cons_acf <- acf(cons, plot = F); cons_pcf <- pacf(cons, plot = F)

par(mfrow=c(2,2))
plot(inc_acf);plot(inc_pcf);plot(cons_acf);plot(cons_pcf)

```

We observe that in both cases, the autocorrelation function shows extremely high estimates for each subsequent lag: even for a large number of periods, there is still a large degree of autocorrelation in the data. The partial autocorrelations, however, show a completely different view: there is no partial autocorrelation when controlled for other influences. This means that there is a lot of dependence in the series unconditionally, but conditionally on previous values, there seems to be no correlation. This behavior seems to be consistent with random walk behavior. 

**Part 2**

We use the same function as in the previous question to find the best model (according to BIC), and compute the t-value and conduct a Dickey-Fuller test:

```{r hallohoegaathet, results='hide'}

urcons <- unit_root_test(df4$CONS, order = 12, -1.9393)
urinc <- unit_root_test(df4$INC, order = 12, -1.9393)

unitroots <- rbind(urcons, urinc) %>%
  mutate(stock = str_replace(stock, "df4\\$", ""))

rownames(unitroots) <- NULL

```

```{r table2}
knitr::kable(unitroots)
```

As we can see in the table, the unit root hypothesis is not rejected at a 5% level in both cases. 

**Part 3**

Now, we perform a unit root test on the first differences of both series:

```{r heehallo, results='hide'}
urcons2 <- unit_root_test(diff(df4$CONS), order = 12, -1.9393)
urinc2 <- unit_root_test(diff(df4$INC), order = 12, -1.9393)

unitroots2 <- rbind(urcons2, urinc2) %>%
  mutate(stock = str_replace(stock, "df4\\$", ""))

rownames(unitroots2) <- NULL
```

```{r een_mooi_tabelletje}
knitr::kable(unitroots2)
```

Now, we see that both hypotheses are rejected! We conclude that both of these series are integrated at order $I(1)$, because the hypothesis of a unit root in the first differences is rejected. 

**Part 4**

```{r, results='asis'}

consincreg <- lm(data = df4, 
   formula = CONS ~ INC)

consincreg %>%
  stargazer(header = F, 
            omit.stat = c("adj.rsq", "ser", "f"))
```



```{r cointegration, results='hide'}
plot(consincreg$residuals)

urt2 <- unit_root_test(consincreg$residuals, 
               order = 12, 
               -1.6156)

rownames(urt2) <- NULL
```

```{r}
knitr::kable(urt2)
```

We observe that the hypothesis of a unit root in the residuals is rejected, indicating that the two series are cointegrated: $Cons, Inc \sim CI(1,1)$. Using general-to-specific approach with BIC, a determined number of lags is two.

**Part 5**

As the series are cointegrated, we can estimate an error correction model, incorporate long-run equilibrium information as well as short-run dynamics. The error-correction model is of the following general form:

\begin{align*}
\Delta Y_t = \alpha - \gamma \bar{Z}_{t-1} + \phi_1 \Delta Y_{t-1} + \dots + \phi_p \Delta Y_{t-p} + \\
\beta_0 \Delta X_{t} + \dots + \beta_q \Delta X_{t-q} + u_t
\end{align*}

We first construct the necessary data matrix in Python:

```{python}
import pandas as pd 

# p for dep var, q for indep var
def generate_datamatrix(p, q, y, x):
  # dependent variable
  fdy = r.df4[str(y)] - r.df4[str(y)].shift(1)
  #ytm1
  ytm1 = r.df4[str(y)].shift(1)
  #xtm1
  xtm1 = r.df4[str(x)].shift(1)
  
  d = {}
  
  d['xt'] = r.df4[str(x)]
  d['xtm1'] = xtm1
  # x first differences
  d['dxtm0'] = r.df4[str(x)] - r.df4[str(x)].shift(1)
  # lagged first differences
  for i in range(1, q+1):
    d['dxtm{0}'.format(i)] = d['dxtm0'].shift(i)
  
  # y first differences
  e = {}
  
  e['yt'] = r.df4[str(y)]
  e['fdy'] = fdy
  e['ytm1'] = ytm1
  
  for j in range(1, p+1):
    e['dytm{0}'.format(j)] = e['fdy'].shift(j)
    
  # put everything in a dataframe
  dfx = pd.DataFrame.from_dict(d)
  dfy = pd.DataFrame.from_dict(e)
  
  df_out = pd.concat([dfx.reset_index(drop=True), dfy], axis=1)
  
  return df_out

data_ecm = generate_datamatrix(4, 4, 'CONS', 'INC')

```

Now, we implement the general-to-specific procedure in R. First, we estimate the residuals $\bar{Z}_t$, and compute $\bar{Z}_{t-1}$ which we then lag to add as a predictor in the general-to-specific procedure. 

```{r}
zt <- lm(data = py$data_ecm,
   formula = yt ~ xt)

residuals <- zt$residuals

ztm1 <- lag(residuals, 1)
```

Now, we add $\bar{Z}_{t-1}$ to the dataset, and formulate the g2s procedure (again according to the BIC):

```{r, results="hide"}
data_ecm <- py$data_ecm %>%
  mutate(ztm1 = ztm1) %>%
  na.omit()

indep_vars <- py$data_ecm %>%
  colnames()

indep_vars <- indep_vars[grepl("dxt|dyt", indep_vars)]

longformula <- paste("fdy ~ ztm1 + 0 +", paste(indep_vars, collapse = " + "))
# run the stepwise regression - find the best model
null = lm(data = data_ecm, formula = "fdy ~ ztm1 + 0")
full = lm(data = data_ecm, formula = longformula)
  
bestmodel <- step(full,
       scope = list(lower = null, upper = full),
       direction = "backward",
       criterion = "BIC",
       k = log(nrow(data_ecm)))
```

Starting with $p = q = 4$, we get the following model:

```{r, results='asis'}

stargazer(bestmodel, header = F,
          omit.stat = c("adj.rsq", "ser", "f"))
```

_Report and interpret the short-run and long-run multipliers.  Report and interpret the error correction coefficient._

<!-- Lecture 4, slide 78 and onward: error correction model interpretation -->

The error correction coefficient $\gamma$ = `r bestmodel$coefficients[1]` means that if consumption and income are away from their long-run relationship, there is a correction towards the equilibrium value. For example, if there is a positive residual, meaning consumption is too high relative to the income at a given point in time, in the next period, we tend to observe a _decrease_ in consumption to correct for that. 

The short-run multipliers can be derived from the coefficients belonging to $\Delta X_t$ and $\Delta Y_{t-3}$. An (positive) exogenous shock in income causes an increase in consumption (0.19), and the effects of that shock also appear three periods later $\Delta Y_{t-3}$. 

The long-run multipliers can be derived from the equilibrium relationship in the "first-stage" regression:

$\bar{Y} =$  `r zt$coefficients[1]` + `r zt$coefficients[2]` $\cdot \bar{X}$.  

This indicates that people tend to consume a proportion of 66% of their long-run equilibrium income, plus a fixed amount of `r zt$coefficients[1]`, which can (kind of) be interpreted as the cost of living (even if the long-run equilibrium income is 0, they theoretically consume this amount). 

_6.  How strong is the correction to equilibrium?  Is there over-shooting?_

The error correction coefficient $\gamma$ = `r bestmodel$coefficients[1]` is very close to zero. Hence, there is no overshooting $\gamma < -1$, and only a partial error correction. The error correction is also quite slow, indicated by the low magnitude of the coefficient. 

_Do you find evidence of Granger causality?  Justify your answer._

We can formally test for Granger causality by taking a stationary version of both series (in this case, the first differences) and investigate if there is any predictive power from the lagged first differences of $X_t$ (income), conditional on the inclusion of lagged $Y$ (first differences of consumption) variables. To do this, we again use the Python code to generate a datamatrix containing first differences of both variables, and lagged versions. We start with 10 lags:

```{python}
data_granger = generate_datamatrix(10, 10, 'CONS', 'INC')
```

We now use this dataset in R to perform stepwise regression:

```{r, results='hide'}

indep_vars <- py$data_granger %>%
  colnames() 

indep_vars <- indep_vars[grepl("dxtm|dytm", indep_vars)]

longformula <- paste("fdy ~ ", paste(indep_vars, collapse = " + "))
# run the stepwise regression - find the best model
null = lm(data = py$data_granger %>%
            na.omit(), formula = "fdy ~ dytm1")
full = lm(data = py$data_granger %>%
            na.omit(), formula = longformula)
  
bestmodel <- step(full,
       scope = list(lower = null, upper = full),
       direction = "backward",
       criterion = "BIC",
       k = log(nrow(data_ecm)))
```

```{r results ='asis'}
stargazer(bestmodel, 
          omit.stat = c("f", "ser", "adj.rsq"),
          header = FALSE)
```

We can find very clear evidence of Income Granger-causing Consumption: there is at least 1 (in the case) two lagged variables that are significant: $\Delta X_{t-4}$ and $\Delta X_{t-8}$. This is probably because there is a seasonality effect. In any case, if at least one of the lags is significant, then $\Delta X$ granger-causes $\Delta Y$. The p-values are small enough, so that we also don't have two worry about type I-errors while making this conclusion. 

## Question 5

First, we import and plot the data:

```{python}
import dateutil
import matplotlib.pyplot as plt

coviddixneuf = pd.read_csv("./data/data_assign_p5.csv", skiprows=2)
coviddixneuf['Date'] = coviddixneuf['Date'].apply(dateutil.parser.parse, dayfirst=True)

plt.plot(coviddixneuf['Date'], coviddixneuf['New Cases'])
plt.title("Covid Cases over Time (America)")
```

We observe that both the mean and the variance tend to vary over time. This means we cannot use models of the ARMA(p,q) class to model the process, because, whereas the conditional mean changes over time and can thus accommodate the characteristics of this process, the variance does not. To model this behavior, we might resort to the GARCH-class of models. These models can be applied to stationary data series, and fit an ARMA model with conditional heteroskedasticity. 

```{r}
library(fGarch)

covid19 <- py$coviddixneuf %>%
  janitor::clean_names() %>%
  arrange(date)

garchmodel <- fGarch::garchFit(formula = ~ arma(2, 1) + garch(1,1), 
                               data = covid19$new_cases,
                               trace = F)

eval <- data.frame(date = covid19$date[1:length(covid19$date)], 
                   actual = covid19$new_cases, 
                   fitted = garchmodel@fitted)

# first differences
eval %>%
  pivot_longer(c(actual, fitted)) %>%
  ggplot(aes(x = date, y = value, group = name, color = name)) + geom_line()

test <- predict(garchmodel, n.ahead = 300)
```

