---
title: | 
  | Econometrics III
  | Assignment Part 3, 4, 5
  | Tinbergen Insitute
author: |
  | Stanislav Avdeev \hspace{3em} Bas Machielsen 
  | 590050sa \hspace{5em} 590049bm 
  | stnavdeev@gmail.com \hspace{2em} 590049bm@eur.nl
date: \today
output: 
  pdf_document: 
    latex_engine: lualatex
---

```{r setup, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(echo = TRUE,  # Если мы не хотим код-чанков, можем менять на echo = FALSE
                      warning = FALSE,
                      message = FALSE,
                      size = "small",
                      out.width = "300pt", out.height = "200pt", 
                      fig.align = "center")

library(rio); library(tidyverse)
library(dynlm) ; library(stargazer)
library(normtest); library(reticulate)
library(AER)

```

<!-- I created a file .Renviron, where the python distribution is located on my system. You can find that by opening a terminal, and entering `$ which -a python python3`. Then, in RStudio, use `usethis::edit_r_environ()`, and add `RETICULATE_PYTHON="/Users/basmachielsen/opt/anaconda3/bin/python"` (or your directory) on a new line to the file. In this way, we can seamlessly interchange R and Python code chunks. Restart RStudio, and then everything is ready to go: --> 

## Question 1

**Part 1**:

First, we plot the two time series:

```{r plot}
df3 <- readr::read_csv("./data/data_assign_p3.csv")

twostocks <- c("apple", "netflix")

df3_twostocks <- df3 %>%
    janitor::clean_names() %>%
    pivot_longer(-date) %>%
    filter(is.element(name, twostocks)) %>%
    mutate(date = lubridate::dmy(date)) 

df3_twostocks %>%
    ggplot(aes(x = date, y = value, 
               group = name, color = name)) + geom_line()
```

Then, we show the acf and pacf-functions:

```{r autocor}

n1 <- df3$NETFLIX %>%
    acf(lag.max = 12, plot = F)

n2 <- df3$NETFLIX %>%
    pacf(lag.max = 12, plot = F)

a1 <- df3$APPLE %>%
    acf(lag.max = 12, plot = F)

a2 <- df3$APPLE %>%
    pacf(lag.max = 12, plot = F)

par(mfrow=c(2,2))

plot(n1, main = "Netflix");plot(n2, main = "Netflix")
plot(a1, main = "Apple"); plot(a2, main = "Apple")
     
```

The ACF's tell us that the stock price is highly dependent on the past stock price, and this dependence decays only very slowly: even the 50 or 100-period lag still shows significant autocorrelation. 

\clearpage

**Part 2**

We now implement a general to specific unit root test function:

```{r g2s function}
unit_root_test <- function(column, order, critical_value){
  
  # Make the dataset
  series <- ts(column) 
  first_differences <- diff(series, differences = 1)
  laggedvar <- stats::lag(series, -1)
  
  #other lagged first differences, delta x_{t-1}, .., delta x_{t-p+1}
  lagged_fds <- list()
  
  for(i in 1:(order-1)){
    
    lagged_fds[[i]] <- stats::lag(first_differences, k = -i)
    
  }
  
  df <- cbind(first_differences, laggedvar, purrr::reduce(lagged_fds, cbind)) %>%
    as_tibble()
 
  colnames(df) <- c("dxt", "xtm1", paste("dxtm", 1:(order-1), sep = ""))
  
  df <- df %>%
    na.omit()
  
  # run the stepwise regression - find the best model
  null = lm(data = df, formula = "dxt ~ xtm1")
  full = lm(data = df, formula = paste("dxt ~ xtm1 +", 
                                       paste(paste("dxtm", 1:(order-1), sep = ""), 
                                             collapse = ' + '),
                                       collapse = " ")
  )
  
  bestmodel <- step(full,
       scope = list(lower = null, upper = full),
       direction = "backward",
       criterion = "BIC",
       k = log(nrow(df)))
  
  # perform the unit root test (MacKinnon, 2010)
  b_critical <- critical_value
  
  t_value <- bestmodel %>%
    summary() %>%
    .$coefficients %>%
    .[,3] %>%
    .["xtm1"]
  
  significant = abs(t_value) > abs(b_critical)
    
  data.frame(stock = deparse(substitute(column)),
             best_model = as.character(bestmodel$call[2]), 
             t_value = t_value, 
             sig = significant)
}


```


```{r, message = FALSE, warning = FALSE, results = "hide"}
summary_tests <- list()

for(i in 1:length(colnames(df3[,-1]))){
  df <- unit_root_test(df3[, i+1], 12, -1.6156)
  summary_tests[[i]] <- df %>%
    mutate(name = colnames(df3[,-1][i]))
}

summary_tests <- summary_tests %>%
  purrr::reduce(rbind) %>%
  select(-stock)

rownames(summary_tests) <- NULL
```

```{r table}
knitr::kable(summary_tests)
```

The test seems to be significant for a number of stocks, indicating that for these stocks, the null hypothesis of a unit root is rejected in favor of stationarity. With a 10% $\alpha$-level, we expect to see a type I-error (rejecting the null while it is true) about one tenth of the time for every test. This means that the probability of having at least 1 type-I error is very large: $(1 - 0.9^{10})$ = `r 1-0.9^10`. It would be better to use some kind of Bonferroni correction to correct for these compounding type I-errors, but alternatively, we could also lower the $\alpha$-level. 

**Part 3**

The forecast $\mathbb{E}[P_{t+1} | D_t] = \mathbb{E}[P_{t+1} | P_t] = \mathbb{E}[P_t + \epsilon_t] = P_t$. Similarly, the forecast $\mathbb{E}[P_{t+2}] = \mathbb{E}[P_{t+1} + \epsilon_{t+1}] = \mathbb{E}[P_{t+1}] = P_t$. Generalizing this pattern, the forecast for $P_{t+h} = P_t$. The variance of the forecast is derived using the distribution:

\begin{align*}
P_{t+1} &= P_t + \epsilon_t \Rightarrow P_{t+1} | P_t \sim N(P_t, \sigma^2) \\
P_{t+2} &= P_{t+1} + \epsilon_{t+1} = \\
&= P_t + \epsilon_t + \epsilon_{t+1} \Rightarrow P_{t+2} | P_t \sim N(P_t, 2 \sigma^2) \\
\end{align*}

Generalizing this pattern, we can see that $\text{Var}(P_{t+h}) = h \cdot \sigma^2$. Hence, we can implement our forecasts in the following way:

```{r}
#forecast code
p_tplush <- df3 %>%
  slice_tail(n=1) %>%
  select(c("APPLE", "MICROSOFT")) 

var_apple <- lm(data = df3 %>%
     select("APPLE"), 
   formula = APPLE ~ lag(APPLE, 1)) %>%
  .$residuals %>%
  var()

var_microsoft <- lm(data = df3 %>%
     select("MICROSOFT"), 
   formula = MICROSOFT ~ lag(MICROSOFT, 1)) %>%
  .$residuals %>%
  var()

forecasts_apple <- data.frame(time = 1:5, 
                              value = rep(p_tplush %>%
                                            pull(1), 5),
                              var = var_apple * 1:5)

forecasts_microsoft <- data.frame(time = 1:5, 
                                  value = rep(p_tplush %>%
                                               pull(2), 5),
                                  var = var_microsoft * 1:5)

forecasts_apple %>%
  ggplot(aes(x = time)) + 
  geom_line(aes(y = value)) +
  geom_line(aes(y = value + 1.96*sqrt(var)), lty = "dashed") +
  geom_line(aes(y = value - 1.96*sqrt(var)), lty = "dashed")

forecasts_microsoft %>%
  ggplot(aes(x = time)) + 
  geom_line(aes(y = value)) +
  geom_line(aes(y = value + 1.96*sqrt(var)), lty = "dashed") +
  geom_line(aes(y = value - 1.96*sqrt(var)), lty = "dashed")

```

Hence, there is no investment advice that we can give: the value is predicted to remain constant, and for all predictions, the probability of an increase in stock price equals the probability of a decrease in stock price. The expected value of any investment strategy is 0, and the value is neither expected to increase, nor to decrease. 

**Part 4**

_Do  you  find  a  statistically  significant  contemporaneous  relation  between  Microsoft  and Exxon  Mobile  stock  prices?_

```{r, results='asis'}
lm(df3, formula = MICROSOFT ~ EXXON_MOBIL) %>%
  stargazer(header = F, omit.stat = c("adj.rsq", "ser", "f"))

```

_Do  you  agree  that  changes  in  Microsoft  stock  prices  are largely explained by fluctuations in the stock price of Exxon Mobile?_

We do not agree that changes in Microsoft stock prices are largely explained by fluctuations in the stock price of Exxon Mobile. It is likely that these results are driven by a shared stochastic trend in both of variables, in other words, the variables might be cointegrated. It can be shown that if two variables share a stochastic trend, the t-value of the estimated coefficient tends to infinity, and the probability of obtaining statistical significance to 1, even under the assumption of completely unrelated trends. 

## Question 4

**Part 1**

```{r}
df4 <- read_csv("data/data_assign_p4.csv")

df4$obs <- ts(df4$obs, 
              frequency = 4,
              start = c(1988, 1))

p_cons <- ggplot(data = df4, aes(x = as.Date(obs), y = CONS)) + 
  geom_line() + 
  scale_x_date(date_labels = "%b %Y")

p_inc <- ggplot(data = df4, aes(x = as.Date(obs), y = INC)) + 
  geom_line() + 
  scale_x_date(date_labels = "%b %Y")

cowplot::plot_grid(p_cons, p_inc)

inc <- ts(df4$INC,
   frequency = 4,
   start = c(1988, 1)) 

cons <- ts(df4$CONS,
   frequency = 4,
   start = c(1988, 1)) 

inc_acf <- acf(inc, plot = F); inc_pcf <- pacf(inc, plot = F)
cons_acf <- acf(cons, plot = F); cons_pcf <- pacf(cons, plot = F)

par(mfrow=c(2,2))
plot(inc_acf);plot(inc_pcf);plot(cons_acf);plot(cons_pcf)

```

We observe that in both cases, the autocorrelation function shows extremely high estimates for each subsequent lag: even for a large number of periods, there is still a large degree of autocorrelation in the data. The partial autocorrelations, however, show a completely different view: there is no partial autocorrelation when controlled for other influences. This means that there is a lot of dependence in the series unconditionally, but conditionally on previous values, there seems to be no correlation. This behavior seems to be consistent with random walk behavior. 

**Part 2**

We use the same function as in the previous question to find the best model (according to BIC), and compute the t-value and conduct a Dickey-Fuller test:

```{r hallohoegaathet, results='hide'}

urcons <- unit_root_test(df4$CONS, order = 12, -1.9393)
urinc <- unit_root_test(df4$INC, order = 12, -1.9393)

unitroots <- rbind(urcons, urinc) %>%
  mutate(stock = str_replace(stock, "df4\\$", ""))

rownames(unitroots) <- NULL

```

```{r table2}
knitr::kable(unitroots)
```

As we can see in the table, the unit root hypothesis is not rejected at a 5% level in both cases. 

**Part 3**

Now, we perform a unit root test on the first differences of both series:

```{r heehallo, results='hide'}
urcons2 <- unit_root_test(diff(df4$CONS), order = 12, -1.9393)
urinc2 <- unit_root_test(diff(df4$INC), order = 12, -1.9393)

unitroots2 <- rbind(urcons2, urinc2) %>%
  mutate(stock = str_replace(stock, "df4\\$", ""))

rownames(unitroots2) <- NULL
```

```{r een_mooi_tabelletje}
knitr::kable(unitroots2)
```

Now, we see that both hypotheses are rejected! We conclude that both of these series are integrated at order $I(1)$, because the hypothesis of a unit root in the first differences is rejected. 

**Part 4**

Assuming both series are I(1), test for cointegration betweenconsumptionandincomebyregressingconsumptiononincomeand performing a unit-root test on the residuals.  Reportthe  estimated  regression  coefficients.  

Plot  the  regression  residuals.   Use  the  SchwartzInformation Criterion (SIC) to determine the number of ADF lags in your unit-root residualtest.  Report the cointegration test statistic.  Do you reject cointegration?

```{r henk}

consincreg <- lm(data = df4, 
   formula = CONS ~ INC)

consincreg %>%
  stargazer(header = F, 
            omit.stat = c("adj.rsq", "ser", "f"))
```



```{r cointegration, results='hide'}

unit_root_test(consincreg$residuals, 
               order = 12, 
               -1.6156)

```

We observe that the hypothesis of a unit root in the residuals is rejected, indicating that the two series are cointegrated: $Cons, Inc \sim CI(1,1)$. 

**Part 5**

As the series are cointegrated, we can estimate an error correction model, incorporate long-run equilibrium information as well as short-run dynamics. The error-correction model is of the following general form:

$$

$$
